Developing a preprint culture in biology

### 2013-02-25 ms.tex
rev following Dr. Gross' email

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{The Case for Preprints in Biology}
}
% Insert Author names, affiliations and corresponding author email.
\\
Philippe Desjardins-Proulx$^{1, 2,\ast}$, 
Ethan P. White$^{3}$, 
Joel J. Adamson$^{4}$,
Karthik Ram$^{5}$,
Timoth\'ee Poisot$^{1,2,6}$,
Dominique Gravel$^{1,2}$
\\
\bf{1} Theoretical Ecosystem Ecology laboratory, Universit\'e du Qu\'ebec \`a Rimouski, Quebec, Canada.
\\
\bf{2} Quebec Center for Biodiversity Science, McGill University, Quebec, Canada.
\\
\bf{3} Departement of Biology, Utah State University, Utah, United-States of America.
\\
\bf{4} Ecology, Evolution and Organismic Biology, University of North Carolina at Chapel Hill, North Carolina, United-States of America.
\\
\bf{5} Environmental Science, Policy, and Management, University of California, Berkeley, California, United-States of America.
\\
\bf{6} International Network for Next-Generation Ecology.
\\
$\ast$ E-mail: philippe.d.proulx@gmail.com
\end{flushleft}

@@

\section*{Current offerings}

We briefly discuss several options to submit preprints to open servers:
arXiv.org, figshare, and the upcoming PeerJ and F1000Research.

@@

arXiv (\url{http://arxiv.org/}) is the most widely-used preprint server today,
and its use is almost universal in some branches of mathematics and physics.
arXiv provides a reliable citation system for all eprints. Physicist Paul
Ginsparg created arXiv in 1991 for theoretical high-energy physicists to
communicate preprints via email and ftp, and soon thereafter adopted the newly
created world-wide web \cite{jackson2002preprints}.  arXiv now receives over 7
000 submissions per month (\url{http://arxiv.org/show_monthly_submissions}) and
divides its submissions into subcategories of physics, mathematics, computer
science, quantitative biology, finance and statistics.  The quantitative biology
category includes subcategories for Populations and Evolution, Quantitative
Methods and other categories that may be of interest to biologists.

Unlike the other options presented here, arXiv has a system of moderators 
and endorsors. At least one author of a paper must be an endorser that has 
previously submitted a paper or has received permission to submit. 
Moderators have the power to change the classification of a paper. 
arXiv is now administered by the Cornell University Libraries, with funding 
coming from voluntary pledges by academic institutions along with matching 
funds from the Simons Foundation \cite{arxiv_future}. This approach to 
financing is one of the numerous measures that arXiv takes to ensure that 
the repository will remain permanently available and submissions will be 
readable.

@@

figshare (\href{http://figshare.com}{http://figshare.com}) is an open server
allowing scientists to submit any research output: manuscript, figures,
datasets, videos, theses, presentations, and so on. There are no rules to limit
what constitutes a research output and, unlike arXiv, there is no endorser
system. All figshare content has a unique digital object identifier (DOI) like
any journal article, thus offering a permanent and stable link to the content.
A flexible tag system is used to classify each item. Comments can be made on all
content allowing for centralized discussion related to the material.

Unlike arXiv, figshare is not limited to quantitative sciences. arXiv.org has
sections on quantitative biology but might not be appropriate for
non-quantitative work. With its flexible approach to preprints, figshare offers
an alternative to arXiv for empirical biologists. Furthermore, by allowing all
types of content, figshare also provides an archive for early results
(\emph{e.g.}: figures, lab presentations).

@@

PeerJ (\href{https://peerj.com/}{https://peerj.com/}) is a new open access
publisher that combines both a preprint server and a peer reviewed journal.  It
is focused on the biological and medical sciences, which may help overcome the
perception that preprints do not have a place in biology. Like figshare, PeerJ
allows commenting on posted preprints, improving the potential for
pre-publication dialog. In addition, preprints can be made private if the
authors choose, and shared only with selected colleagues. While this reduces
many of the benefits of preprints described above, it may allow some researchers
who would not otherwise post preprints to begin exploring the possibility in a
manner appropriate to their current circumstances.

In contrast to other preprint servers, users cannot post unlimited public
preprints for free. One preprint per year can be posted for free, with a onetime
(\emph{i.e.} lifetime) fee of 99 dollars required to allow an author to post
unlimited public preprints. The preprint server is not tied to the journal, so
preprints can be posted regardless of where they will eventually be submitted
for publication.

@@

F1000Research is not a public preprint server like the previous three servers.
Whereas arXiv, figshare, and PeerJ offer an option to submit a manuscript
without having it reviewed, papers submitted to F1000Research will eventually be
reviewed. Thus, F1000Research offers a hybrid model with publicly available
manuscripts at time of submission and standard peer-reviews that occur as part
of the submission process. Manuscripts are considered ``accepted'' and will only
be indexed after two positive referee responses. F1000Research works closely
with data providers to integrate raw data to the paper. For instance, upon
submitting a paper, authors are asked to upload their data, which are then
integrated in \emph{e.g.} figshare widgets, the DOI of which are given in the
paper when the data are first mentioned.  By integrating data to the paper,
F1000Research is working to make science more reproducible and open.

@@

This manuscript was developed entirely as an open project on GitHub. GitHub is
one of several hosting services for collaborative development using the Git
version control system (VCS).  Git is a decentralized revision control system
created by Linus Torvalds and is used primarily to develop software, including
the Linux kernel. Git provides powerful features that allow numerous
contributers to work asynchronously on the same project, often in parallel
branches, all of which can be effortlessly merged and version controlled.  While
Git is created primarily for software development, where the use of version
control systems is standard \cite{aru12}, it is ideal for academic research
since it provides a way to collaborate on every step of the manuscript
development process. For example, during the development of this manuscript,
each author would clone the project (\emph{i.e.} make a personal copy), modify
it, and then merge their changes into a master branch. This takes the preprint
process to an new level, where the entire writing process is open from the
beginning.

@@

@@

>>>>

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{The Case for Open Preprints in Biology}
}
% Insert Author names, affiliations and corresponding author email.
\\
Philippe Desjardins-Proulx$^{1, 2,\ast}$, 
Ethan P. White$^{3}$, 
Joel J. Adamson$^{4}$,
Karthik Ram$^{5}$,
Timoth\'ee Poisot$^{1,2,6}$,
Dominique Gravel$^{1,2}$
\\
\bf{1} Theoretical Ecosystem Ecology laboratory, Universit\'e du Qu\'ebec \`a Rimouski, Quebec, Canada.
\\
\bf{2} Quebec Center for Biodiversity Science, McGill University, Quebec, Canada.
\\
\bf{3} Departement of Biology, Utah State University, Utah, United-States of America.
\\
\bf{4} Ecology, Evolution and Organismic Biology, University of North Carolina at Chapel Hill, North Carolina, United-States of America.
\\
\bf{5} Environmental Science, Policy, and Management, University of California, Berkeley, California, United-States of America.
\\
\bf{6} International Network for Next-Generation Ecology.
\\
$\ast$ E-mail: philippe.d.proulx@gmail.com
\end{flushleft}

@@

\section*{Box: Preprint Server Roundup}

% Table 2 should start here.

@@

arXiv (\url{http://arxiv.org/}) is the most widely-used preprint server today,
and its use is almost universal in some branches of mathematics and physics.
arXiv has a system of moderators and endorsers. At least one author of a paper
must be an endorser that has previously submitted a paper or has received
permission to submit.  Moderators have the power to change the classification of
a manuscript.

@@

figshare (\href{http://figshare.com}{http://figshare.com}) is an open server
allowing scientists to submit any research output: manuscript, figures,
datasets, videos, theses, presentations, and so on. There are no rules to limit
what constitutes a research output and, unlike arXiv, there is no endorser
system. A flexible tag system is used to classify each item.

@@

PeerJ (\href{https://peerj.com/}{https://peerj.com/}) is a new commercial open
access publisher that combines both a preprint server and a peer reviewed
journal. Preprints can be made private and shared only with selected colleagues.
Only one preprint per year can be posted for free, with a onetime (\emph{i.e.}
lifetime) fee required to allow an author to post unlimited public preprints.
Preprints can be posted to PeerJ regardless of where they will eventually be
submitted for publication.

@@

Whereas arXiv, figshare, and PeerJ offer an option to submit a manuscript
without having it reviewed, papers submitted to F1000Research will eventually be
reviewed. Thus, F1000Research offers a hybrid model with publicly available
manuscripts at time of submission and standard peer-reviews that occur as part
of the submission process. Manuscripts are considered ``accepted'' and will only
be indexed after two positive referee responses.

@@

This manuscript was developed entirely as an open project on GitHub. GitHub is
one of several hosting services for collaborative development using the Git
version control system (VCS). It allows numerous contributers to work
asynchronously on the same project, often in parallel branches, all of which can
be effortlessly merged and version controlled.  Git is primarily used for
software development \cite{aru12} but it provides a powerful tool way to
collaborate on every step of the manuscript development process.

@@

\section*{Competing interests}

DG is an associate falculty of F1000, EW is an academic editor for PeerJ, and TP
is an advisor for figshare.

@@

\begin{table}[!ht]
    \caption{\bf{Popular options for preprints}}}
    \begin{tabular}{|l|cccccccc|}
    \hline
    Website               & Free & Comments & Private preprints & Moderated & Peer-reviewed & DOI & Version-Control & Other Content\\
    \hline
    arXiv.org             & Yes & No & No & Yes & No & No & No & No\\
    figshare.com          & Yes & Yes & No & No & No & Yes & No & Yes\\
    peerj.com             & No & Yes & Yes & No & No & Yes & No & No\\
    f1000research.com     & No & Yes & No & Yes & Yes & Yes & No & No\\
    github.com            & Yes & Yes & Yes & No & No & No & Yes & Yes\\
    \hline
    \end{tabular}
    \label{table:box1}
\end{table}
